### 2016-12-04 nomen.tex

%% LyX 2.1.5 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[english]{article}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage{nomencl}
% the following is useful when we have the old nomencl.sty package
\providecommand{\printnomenclature}{\printglossary}
\providecommand{\makenomenclature}{\makeglossary}
\makenomenclature
\usepackage{babel}
\begin{document}

\title{Nomenclature}

\maketitle
\nomenclature{WBN}{Wireless Butterfly Network}\nomenclature{WNC}{Wireless Network Coding}\nomenclature{HW}{Hardware}\nomenclature{SNR}{Signal to Noise Ratio}\nomenclature{SC}{Superposition Coding}\nomenclature{HSI}{Hierarchical Side Information}\nomenclature{DEM}{Demodulator}\nomenclature{IC}{Interference Cancellation}\nomenclature{MA}{Multiple Access}\nomenclature{AWGN}{Aditive White Gaussian Channel}\nomenclature{ASK}{Amplitude Shift Keying}\nomenclature{QAM}{Quadrature Amplitude Modulation}\nomenclature{SER}{Symbol Error Rate}\nomenclature{FER}{Frame Error Rate}\nomenclature{SDR}{Software Defined Radio}\nomenclature{USRP}{Universal Software Radio Peripherals}\nomenclature{UHD}{USRP Hardware Driver }\nomenclature{CP}{Cyclic Prefix}\nomenclature{ODFD}{Orthogonal Frequency Division Multiplexing}\nomenclature{DC}{Direct Current}\nomenclature{MIMO}{Multiple Input Multiple Output}\nomenclature{QPSK}{Quadrature Phase Shift Keying}\nomenclature{UDP}{User Datagram Protocol}\nomenclature{BC}{Broadcast Channel}\nomenclature{MAC}{Multiple Access Channel}\nomenclature{CC}{Constellation Constrained}\settowidth{\nomlabelwidth}{AWGN}
\printnomenclature{}
\end{document}