### 2017-05-30 response_letter/response_letter.tex
Few new lines in response letter

%% LyX 2.1.3 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
%\RequirePackage[loading]{tracefnt}
\documentclass[10pt,a4paper,titlepage,conference]{article}
\usepackage[T1]{fontenc}
%\usepackage[latin9]{inputenc}
%\usepackage[utf8]{inputenc}
%\usepackage[english]{babel} 
\usepackage{units}
\usepackage{bm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}

\makeatletter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
\special{papersize=\the\paperwidth,\the\paperheight}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
%\documentclass[8pt,titlepage,a4paper]{article}%{/usr/share/texmf-dist/tex/latex/IEEEtran/IEEEtran}
%\documentclass[11pt,titlepage,a4paper]{/usr/share/texmf-dist/tex/latex/IEEEtran/IEEEtran}
%\usepackage[latin2]{inputenc}
%\usepackage{times}
%\renewcommand\baselinestretch{1.5}  % 
\usepackage{color}
\usepackage{wrapfig}
\usepackage{epsfig}
%\usepackage{CoverPage}
\usepackage{multirow}
\usepackage{subfigure}
\usepackage{holtpolt}
\usepackage{turnstile}
\usepackage{tikz}
\usepackage{bm}
\usepackage{cite}
\usepackage[multiple]{footmisc}
%\usepackage{calrsfs}
\usetikzlibrary[arrows,decorations.pathmorphing,backgrounds,positioning,fit,petri]

\def\bo#1{{\bf{#1}}}
\def\E{{\rm{E}}}
\def\grebo#1{\mbox{\boldmath$#1$}}
\def\bold#1{\mbox{\boldmath$#1$}}
\def\conv{*}

\def\date#1{\def\@date{#1}}
\def\version#1{\def\@version{#1}}
\def\author#1{\def\@author{#1}}
\def\title#1{\def\@title{#1}}
\def\assignment#1{\def\@assignment{{\it #1}}}
\newcommand{\var}{\rm{var}}

\newcommand{\unfootnote}[1]{
  \renewcommand{\@makefnmark}{}
  \footnotetext{#1}
  \renewcommand{\@makefnmark}{\mbox{$^{\@thefnmark}$}}
}

% \papertitle[title of the paper][Author(s)][Conference type][date + place] 
%\graphicspath{{fig/}}

\begin{document}
\title{Response Letter for EURASIP Journal on Wireless Communications and Networking 
JWCN-D-17-00004 \vskip 1ex
Design of Adaptive Constellations and Error Protection Coding for
Wireless Network Coding in 5-node Butterfly Networks}
\author{Pavel Prochazka, Tomas Uricar, David Halls and Jan Sykora}
\date{\today}
\maketitle
\newpage

We would like to thank all the reviewers for their valuable comments and suggestions. We first reply
to the most common and important questions and then we answer the individual questions and comments
individually.

\section{Common Questions and Corresponding Answers}
\begin{itemize}
  \item CQ1: \emph{Lacking novelty compared to \cite{Prochazka-Uricar-Halls-Sykora_2015-ICC}.}
  \item A1: Compared with the conference version \cite{Prochazka-Uricar-Halls-Sykora_2015-ICC}, the major contributions and differences of this submitted manuscript are summarized as follows:
  \begin{enumerate} 
  \item We introduce additional performance analysis results for constellations in an uncoded system and an in-depth discussion of the observed results.
  \item We extend the section dealing with the adaptive constellation design in an uncoded system and we supplement additional HW evaluation results for the adaptive system in the whole range of observed channel SNRs. Again a close agreement between between the analytic and measurement results is shown.
  \item We show how the channel coding can be integrated into the proposed constellation design, including the cut-set bound analysis of proper source transmission rates, which provides a hint for the setup of all channel encoders' rates in the system.
  \item We evaluate numerically the performance of the resulting adaptive modulation-coding scheme
  for a wide range of channel conditions in the WBN and we emphasise the performance gains with
    respect to the uncoded system.
  \item We perform a simple robustness analysis, demonstrating that the proposed modulation and coding strategy is viable even in the case when the real world condition induce some deviations from the mathematical system model assumed in the paper.
  \item We conducted a comparison with the state of the art reference scenarios (traditional network
  coding and routing) for the proposed channel coded adaptive constellation design.
  \end{enumerate}
\end{itemize}

\begin{itemize}
  \item CQ2: \emph{Better reference scenario.}
  \item A2: We definitely agree that the zero throughput uncoded QPSK reference scenario is not a very good
  to demonstrate the performance gain of the proposed constellation design. 
  Although, the 2-step scenario does not offer other state of the art solution, we
  decided to consider a comparison with traditional network coding and routing strategies. These
  strategies require 3 steps and therefore some attention is required to provide a fair comparison
  with the proposed 2-step framework. We added a completely new section 7.4 into the manuscript
  dedicated to this comparison  with the conventional techniques.

  Within this new section, we shown that
  the HW implementation of the proposed adaptive constellation design outperforms the reference
  scenarios even with the relaxed assumption (pre-rotation off). The good performance of the proposed
  design without pre-rotation is very important, because the
  pre-rotation stands for an implementation challenge requiring the feedback channel. Finally we shown that the
  performance of the proposed framework is even above the corresponding 3-step reference theoretical limits.
\end{itemize}

\begin{itemize}
  \item CQ3: \emph{Typos and Formatting Issues}
  \item A3: The manuscript was carefully proofread by native speaker for typos. Regarding the
  formatting issues, we searched and corrected the issues we found. Nevertheless, we should stress
  that the manuscript will be transformed to the EURASIP template prior publishing and there will
  be time for final tuning of these issues.
\end{itemize}

\begin{itemize}
  \item CQ4: \emph{Generalization of the proposed Adaptive Constellation Design}
  \item A4: The proposed constellation design is indeed suited for the butterfly network. Regarding
  the generalization, we can take the idea beyond the constellation design also for other
  topologies. For instance, the design of the basic part only is relevant for the two-way relay
  channel (actually)
\end{itemize}

\subsubsection*{Reviewer 1}
\begin{itemize}
  \item Q: \emph{The weak aspect of this paper is that the current paper is a small variation of a
  previous paper by the authors. The authors should further highlight the difference of this paper.}
  \item A: See A1
\end{itemize}

\begin{itemize}
  \item \emph{The authors should edit the paper more carefully. For example, in Figure 5, the
  references are missing. Please go over the whole the paper and make sure similar issues do not
  exist}
  \item A: See A3
\end{itemize}

\subsubsection*{Reviewer 2}
\begin{itemize}
  \item Q: \emph{Since this work builds upon previous work on the application of wireless network
  coding to the WBN, novelty is slightly lacking}
  \item A: see A1
\end{itemize}

\begin{itemize}
  \item Q: \emph{The proposed scheme appears to be quite specific to the WBN, I wonder if it is
  possible to generalize to other similar networks where linear codes yield a good performance.}
  \item A: See A4.  We should stress that the proposed constellation design covers both
  wireless network coding and processing the signal from multiple stages. Join generalization of
  both techniques in the way we did would be extremely challenging, because of time 
  synchronization, pre-rotation, etc. Also note that our approach is basically a 
  practical instance of a well known
  theoretical principle with application of WNC to approximately equal power levels and
  interference cancellation on significantly different power levels. So that its generalization in
  theoretical level is already known. 
\end{itemize}

\begin{itemize}
  \item Q: \emph{The authors should also provide more insight or intuition on the design of the
  constellation and explain why the challenges to apply to other linear networks.}
  \item A: The intuition beyond the design could be partly found in \cite{Prochazka-Uricar-Halls-Sykora_2015-ICC}.  
  Particularly, the idea is to consider bit streams that are divided among individual power levels.
  We need to define 1) the scale of individual levels, 2) the constellation element for all levels.
  Obviously, this design would differ for basic and superposed streams:
  \begin{enumerate}
    \item In case of superposed stream, we need to have both source streams decodable and thus we
    need some kind of orthogonality. This orthogonality is achieved by considering real and
    imaginary axis of the constellation space. 
    \item
  \end{enumerate}
  Our solution considers the simplest possible constellation element (BPSK at given level)
\end{itemize}

\begin{itemize}
  \item Q: \emph{It might be useful to the reader to add a figure for the butterfly network in the
  setting of algebraic network coding for comparison.}
  \item A: see A2 and the new section 7.4 in the manuscript
\end{itemize}

\begin{itemize}
  \item Q: \emph{Since real world networks are usually more complicated, I think it would be really interesting if this systematic design can be generalized to all linear network coding schemes (e.g. random linear network coding).}
  \item Q: \emph{If it cannot be easily generalized, it could be helpful to explain what are the
  difficulties in applying it to other coding schemes.}
  \item A: 
\end{itemize}

\begin{itemize}
  \item Q: \emph{Might be helpful if mention beforehand that the nomenclature is given at the end of the paper (For
	  example, HW is throughout the paper and its definition is only given at the end of the paper)}.
  \item A: The nomenclature is placed according to the EURASIP-author's guidelines. The HW
  abbreviation was overlooked - it is now defined also in its first appearance within the text.
\end{itemize}

\subsubsection*{Reviewer 3}
\begin{itemize}
  \item Q: \emph{In figure 6, 7, 8, 9, the reference scheme using QPSK always has a zero throughput,
  which demonstrates that this reference modulation scheme does not work at all in WBN under the
  proposed SNRs. Then the proposed communication scheme is claimed to have performance enhancement
  for sure. However, logically, this only demonstrates that the proposed algorithm gives positive
    throughput in WBN instead of not working. But it is not enough to say it gives a "performance
    enhancement" comparing with a previous working scheme. So it would be better if the author can
    further compare some other modulation schemes which can also give some positive throughput in
    WBN.}
  \item A: See A2. Considering the two time slot communication, there is indeed no state of the art solution
      for joint decoding, that is the constellation design in sources such that the relay would reliably
      jointly resolve data\footnote{We assume uncoded data in this context only.} from both sources. 
      We thus decided to claim the performance of the proposed $N_b=0$, $N_s=2$ compared to the zero
        throughput QPSK as the throughput gain. Nevertheless, the reviewer has point that it is unfair to claim the
        performance of the proposed $N_b=1$, $N_s=1$ constellation as throughput gain compared to
        zero-throughput QPSK, because a conventional wireless network coding with BPSK would give a
        nonzero throughput in this case. We therefore drop this claim from our Figures 6-9 in
        manuscript (check Figure \ref{Fig:Throughput_comparison}).

\end{itemize}

\begin{itemize}
  \item Q: \emph{As far as we know, according to the 3GPP protocol, there is a Channel Quality Index
  (CQI) vs. modulation format table, which suggests the modulation scheme with respect to difference
  SNR range. According to this table, for example, for the SNR varying from 6dB to 20dB, the
  protocol suggests using 16QAM and 64QAM instead of QPSK. In WBN, it could be different from 3GPP,
  but I suggest the author to test some other reference modulation format and compare the
  difference. }
  \item A: The channel quality index is 3-dimensional (MAC, BC and HSI) in the current state. Its
  evaluation can be easily done according to the theoretical analysis. Nevertheless we observed that 
  Although this table would undoubtedly much simplify the implementation, based on our experience,
  we 
\end{itemize}

\begin{itemize}
  \item Q: \emph{Although, in the future work, the authors have mentioned the fading channel case, it would still be better that the case of fading channels or of longer communication distance could be simulated and experimented. AWGN channel seems to be too ideal. If some initial results on the fading channel case can be added into the paper, it would be better.}
  \item A: TODO (possible to play with in the tutorial)
\end{itemize}

\begin{itemize}
  \item Q: \emph{The robustness analysis is important, and the authors have done some simulations on this aspect. However, it would be more appreciated if a hardware experiment can be done on this aspect.}
  \item A: We fully understand and agree that the robustness analysis deserves further deep
      investigation. Nevertheless we provided some basic analysis including the hardware experiment
        within our work (check Figure 17 in manuscript) giving quite promising results. A more detailed investigation is suited rather
        for a standalone work than for this resubmission.
\end{itemize}

\begin{itemize}
  \item Q: \emph{In Fig. 15, the throughput enhancmence for the coding scheme is shown. It is
  interesting to see that the throughput enhancment has some "peak" and "valley" as the γ\_MAC
  increases. If possible, please give some explanation. Or this is just a random instance that
  happens to be shown by the hardware experiment, that is also fine.}
  \item A: The explanation here is quite straightforward, as it follows from the steep
  gradients of the uncoded throughput as function of SNRs shown in Figure 10. On the other hand the
  coded throughput as a function of SNRs is much smoother (Figure 14) and therefore their difference
  induces peeks and valleys
\end{itemize}

\begin{itemize}
  \item Q: \emph{There are also some minor context error or typos found}
  \item A: See A3
\end{itemize}

\section{Major Changes Summary}
\begin{itemize}
  \item New section 7.4 containing a more advanced comparison of the proposed framework with the
  state of the art strategies.
  \item Redrawn Figures (see changes in Figures \ref{Fig:Throughput_comparison} and
  \ref{Fig:AchievedRates}).
\end{itemize}
\begin{figure}
    %\begin{subfigure}[t]{0.5\textwidth}
    \includegraphics[width=0.45\linewidth]{fig/old_Throughput_HSI_XOR_MAC16_BC20_N2.eps}
    \includegraphics[width=0.45\linewidth]{fig/Throughput_HSI_XOR_MAC16_BC20_N2.eps}\\
    \includegraphics[width=0.45\linewidth]{fig/old_Throughput_HSI_XOR_MAC20_BC20_N2.eps}
    \includegraphics[width=0.45\linewidth]{fig/Throughput_HSI_XOR_MAC20_BC20_N2.eps}\\
    \includegraphics[width=0.45\linewidth]{fig/old_Throughput_HSI_XOR_MAC20_BC16_N2.eps}
    \includegraphics[width=0.45\linewidth]{fig/Throughput_HSI_XOR_MAC20_BC16_N2.eps}\\
    \includegraphics[width=0.45\linewidth]{fig/old_Throughput_HSI_XOR_MAC17_BC17_N2.eps}
    \includegraphics[width=0.45\linewidth]{fig/Throughput_HSI_XOR_MAC17_BC17_N2.eps}\\
	\caption{Uncoded throughputs (old versions left, new versions right).}
	\label{Fig:Throughput_comparison}
	%\caption{Old uncoded throughput for $\gamma_{\mathrm{MAC}}=16$,$\gamma_{\mathrm{BC}}=20$.}
    %\end{subfigure} 
    %\begin{subfigure}[t]{0.45\textwidth}
    %\caption{New the uncoded throughput for $\gamma_{\mathrm{MAC}}=16$,$\gamma_{\mathrm{BC}}=20$.}
    %\end{subfigure} 
\end{figure}

\begin{figure}
    %\begin{subfigure}[t]{0.5\textwidth}
    \includegraphics[width=0.45\linewidth]{fig/old_Rate_Regions_XOR_map_BC10_MAC12_HSI8.eps}
    \includegraphics[width=0.45\linewidth]{../fig/Rate_Regions_XOR_map_BC10_MAC12_HSI8.eps}\\
    \includegraphics[width=0.45\linewidth]{fig/old_Rate_Regions_XOR_map_BC5_MAC5_HSI0.eps}
    \includegraphics[width=0.45\linewidth]{../fig/Rate_Regions_XOR_map_BC5_MAC5_HSI0.eps}\\
	\caption{Enlarged fonts -- Figure 13 in the manuscript (old versions left, new versions right).}
	\label{Fig:AchievedRates}
	%\caption{Old uncoded throughput for $\gamma_{\mathrm{MAC}}=16$,$\gamma_{\mathrm{BC}}=20$.}
    %\end{subfigure} 
    %\begin{subfigure}[t]{0.45\textwidth}
    %\caption{New the uncoded throughput for $\gamma_{\mathrm{MAC}}=16$,$\gamma_{\mathrm{BC}}=20$.}
    %\end{subfigure} 
\end{figure}
\bibliographystyle{plain}
\bibliography{bib/Const_design}
\end{document}